---
title: "Advances in Financial Machine Learning: Chapter 2"
subtitle: "Financial Data Structures"
---

# {.tabset .tabset-fade}

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## 2.1 Motivation

* Derive structured datasets from unstructured or minimally structured data
* Using someone else's processed data usually leads to discovering what they already know

## 2.2 Essential Type of Financial Data

* Fundamental Data
  * Regulatory filings and business analytics, reported quarterly
  * Always published with a lag after the official report date (commonly misaligned b/c of this)
  * Commonly backfilled or reinstated, meaning current values were not available at release
  * Extremely structured, low frequency, and accessible = usu. little unexploited value on its own
* Market Data
* Analytics Data
* Alternative Data

## 2.3 Bars

* Bars are the financial term for rows in a dataset
* Can create regular bars using either standard methods or information-driven methods (the latter is rare in literature)
* Standard Bars
  * Time Bars
    * Aggregated in fixed time intervals
    * Most commonly used bars
    * Should be avoided b/c
      * Markets do not process information at a constant rate (i.e. fewer shares are traded at noon that open)
      * Exhibit poor statistical properties (autocorrelation, heteroskedasticity, non-normality of returns, etc.)
  * Tick Bars
    * Aggregate fixed number of ticks (transactions)
    * Closer to Gaussian distribution of returns
  * Volume Bars
    * Aggregate fixed volume of shares
    * Even closer to Gaussian
    * Convenient for many microstructure theories
  * Dollar Bars
    * Sample fixed amount of trade value
    * More stable than other bars in price fluctuations and splits
    
```{r bars}
d <- setDT(tidyquant::tq_get("^GSPC", from = "1990-01-01"))
# Volume bars (use max vol)
max_vol <- max(d$volume)
n_vol <- ceiling(sum(d$volume) / max_vol)
d_vol <- data.table(
  symbol = "^GSPC",
  date = lubridate::Date(n_vol),
  open = numeric(n_vol),
  high = numeric(n_vol),
  low = numeric(n_vol),
  close = numeric(n_vol),
  volume = numeric(n_vol),
  adjusted = numeric(n_vol)
)
vol <- 0
i_vol <- 1L
i0 <- 1L
set(d_vol, i = i0, j = "open", value = d$open[[i0]])
for (i in seq_len(NROW(d))) {
  vol <- vol + d$volume[[i]]
  if (vol >= max_vol) {
    set(
      d_vol, i = i_vol,
      j = c("date", "high", "low", "close", "volume", "adjusted"),
      value = list(d$date[[i]], max(d$high[i0:i]), min(d$low[i0:i]), d$close[[i]], vol, d$adjusted[[i]])
    )
    vol <- 0
    i0 <- i + 1L
    i_vol <- i_vol + 1L
    if (i0 <= NROW(d)) set(d_vol, i = i_vol, j = "open", value = d$open[[i0]])
  }
}
```

## 2.4 Dealing with Multi-Product Series

## 2.5 Sampling Features
